---
title: "p8105_hw2_pm3260"
author: "Pradeeti Mainali"
date: "2024-10-02"
output: github_document
---

```{r setup, echo=FALSE, message=FALSE}
library(tidyverse)
library(readxl)
library(haven)
```

# Problem 1

```{r nyctransit}
nyc_df = read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") |>
  janitor::clean_names() |>
  select(
    line, station_name, station_latitude, station_longitude,
    route1, route2, route3, route4, route5, route6, 
    route7, route8, route9, route10, route11, 
    entrance_type, entry, vending, ada) |>
  mutate(entry_logical = ifelse(entry == "YES", TRUE, FALSE))
```

The NYC Transit data frame contains the variables that describe routes, entry and entrance type, ADA compliance, station name/latitude/longitude, lines, and vending. 

So far, I have cleaned the names to make them all lower case and selected the variables above to remain in the data set.

The dimensions of this data set is 1868 rows and 20 columns. The data is NOT tidy. 

```{r stations, echo=FALSE, results='hide'}
distinct_stations = nyc_df |>
  distinct(station_name, line) |>
  nrow()

distinct_stations
```

There are 465 distinct stations. 

```{r ada, echo=FALSE, results='hide'}
ada_compliant = nyc_df |>
  filter(ada == TRUE) |>
  distinct(station_name, line) |>
  nrow()

ada_compliant
```

There are 84 ADA compliant stations.

```{r vending, echo=FALSE, results='hide'}
no_vending <- nyc_df |>
  filter(vending == "NO") |>
  summarise(proportion = mean(entry_logical, na.rm = TRUE)) |>
  pull(proportion)

no_vending
```

37% of stations
# Problem 2

# Problem 3
